---
title: "2"
subtitle: 'The "tidyverse"'
author: '`r rmdees::title_author()`'
date: '`r rmdees::title_icons_fa5()`'
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    css: ["robot", "css/custom.css"]
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

## Slides

These slides and all workshop materials are available at **github.com/mkearney/moody_rstats**

---

## Agenda

+ Intro to **{tidyverse}**
   - Base R vs "tidyverse"
   - Philosophy
+ Twitter trolls data
   - From FiveThirtyEight
+ Data visualization
   - **{ggplot2}**
+ Data frames
   - **{tibble}**
   - **{readr}**
+ R projects `.Rproj`
   - Creating projects
   - File paths with **{here}**
   - Importing data/file

---

## The "tidyverse"

```{r}
## load tidyverse
library(tidyverse)
```


## Data set
+ [FiveThirtyEight data sets](https://github.com/fivethirtyeight)
   - Kremlin troll tweetcs
```{r eval=FALSE}
## create temp directory
tmpdir <- tempdir()

## download zip archive
usethis::use_course(
  "https://github.com/fivethirtyeight/russian-troll-tweets/archive/master.zip",
  destdir = tmpdir
)

## store vector of the downloaded CSV paths
files <- list.files(
  file.path(tmpdir, "russian-troll-tweets-master"),
  full.names = TRUE, pattern = "\\.csv$"
)

## specify col types to make sure ID variables are character (not numeric)
col_types <- readr::cols(
  external_author_id = readr::col_character(),
  alt_external_id = readr::col_character(),
  tweet_id = readr::col_character()
)

## read all CSV files
d <- purrr::map(files, read_csv, col_types = col_types)

## cleanup files if not already removed
if (any(file.exists(files))) {
  unlink(files)
}
```

+ The `purrr::map()` functions takes an input vector (e.g., `files`) and 
applies a specified function (in this case `readr::read_csv`) to each element. This 
means in the above code, it returned a **list** of data frames `d`, which isn't
very helpful.
+ To collapse a list of similar data frames into a single data frame, use 
`dplyr::bind_rows()`.

```{r}
## d is currently a list of data frames. to collapse it into single data frame
d <- dplyr::bind_rows(d)
```

There's one last thing. The timestamp formatting is a bit weird. The code below
cleans it up and then converts it to an object of case `POSIXct` (date-time).

```{r}
## define function to reformat date
formate_date_char <- function(x) {
  i <- grep("^[1-9]/", x)
  x[i] <- paste0("0", x[i])
  i <- grep("^\\d{2}/\\d{1}/", x)
  x[i] <- sub("/", "/0", x[i])
  i <- grep("201\\d [1-9]:", x)
  x[i] <- sub(" ", " 0", x[i])
  as.POSIXct(x, format = "%m/%d/%Y %H:%M")
}
## reformat the publish_date variable and override previous
d <- d %>%
   dplyr::mutate(publish_date = formate_date_char(publish_date))
```

Now we can preview the data

```{r}
d
```

+ **IF `d` IS NOT A DATA FRAME WITH OVER 2 MILLION ROWS**...or if downloading 
the full data set failed (or it's going too slow), uncomment the lines 
below and run the code to read-in a random sample (N = 100,000) of the data.

```{r}
## read-in a random sample (N = 100,000) of russia-troll-tweets data
#d <- readRDS(
#  here::here("data", "russian-troll-tweets-100k-sample.rds")
#)
#
### preview random sample
#d
```


